A Unified Treatment of Tractability of Approximation Problems Defined on Hilbert Spaces

### 2023-07-17 GeneralizedTractability2023July17.tex
Updates from Overleaf
Fred M2 2023 Mac Updates

\title{\fred{A Unified Treatment of Tractability of Approximation Problems}}

@@

TODO: We should check consistency of notation when using $\max$ or $\min$. Sometimes we use ``$(\ldots)$'' for the argument, sometimes we use ``$\{\ldots\}$''  \fred{Fred will do this.}

@@

The information complexity of the problem $\SOL_d$ is given by the function $\comp:(0,\infty) \times \naturals \to \natzero$, where $\comp(\varepsilon,d)$ is the number of linear functionals required by the best \peter{admissible} algorithm to satisfy the error criterion \eqref{eq:errcrit}.
By definition, the complexity is non-decreasing  as $\varepsilon$ tends to zero.  We also expect it to be non-decreasing as $d$ tends to infinity, since $d$ is typically the number of variables for the functions in $\cf_d$.

For this case where arbitrary linear functionals are allowed, the optimal solution operator is
\[
\APP_d(f,\varepsilon) = \sum_{i=1}^n \peter{\SOL_d(u_{i,d})} \ip[\cf_d]{f}{u_i},
\]
and the information complexity of our linear problem is
\begin{equation}\label{eq:comp_def}
\comp(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\},
\end{equation}
\peter{see \cite{NovWoz08a}.}
\peter{
Strictly speaking, we are considering the \textit{absolute} error criterion in this paper. 
Alternatively, one could also consider the \textit{normalized} error criterion, where the error in \eqref{eq:errcrit} is divided by the so-called \textit{initial error}, which is the error without sampling the function, or, equivalently, the operator norm of $\SOL_d$. In the latter case, it is known (see, e.g., again \cite{NovWoz08a}) that the information complexity is given by 
\[
\comp^{\rm norm}(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\, \lambda_{1, d}\}.
\]
This means that, essentially, in the tractability analysis below, all criteria would be normalized by the first eigenvalue $\lambda_{1,d}$. For the sake of brevity and clarity of notation, we restrict ourselves to the absolute error criterion here, but presumably analogous results hold for the normalized setting. 
}

The question of tractability is one of determining how fast the \peter{information} complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  We want to bound the complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding. \peter{We would like to point out that generalized tractability specified by a function $T$ has been considered before in the literature. In particular, there are several papers by Gnewuch and Wo\'{z}niakowski (see \cite{GW06,GW08,GW09,GneWoz11a} ), and also \cite[Chapter 8]{NovWoz08a} is devoted to this topic. To our best knowledge, however, these references mainly focus on the special case of tensor product problems, which we do not do here, and the conditions for tractability derived there are mostly in terms of properties of the function $T$, and not in terms of sums of (functions of) the eigenvalues of a problem. Hence, we consider the present paper an addition to what is presently known on this subject.}

@@

\kachi{We denote the set of positive integers by $\naturals$ and and the set of non-negative integers by $\naturals_0$.}
In Section \ref{sec:spt}, we give the definitions of the considered tractability notions, in particular (strong) tractability. We consider strong tractability when the information complexity is independent of the dimension, $d$. We then turn to tractability  when the information complexity \peter{may depend on both} $d$ and $\varepsilon^{-1}$. In Section \ref{sec:examples} we provide examples of \peter{tractability functions} for both the algebraic and exponential cases. We also consider other notions of tractability such as quasi-polynomial tractability, and a generalized notion of weak tractability. In the case of generalized weak tractability, we give the necessary and sufficient conditions on the notion of exponential tractability in Section \ref{sec:subh} \peter{[I think I do not quite understand what we mean by the last sentence.]}.

@@

\subsection{Notation and fundamental definitions} \label{sec:not}
Let $T$ be a function given in \eqref{eq:Tspec}.
A tractability function $T$ is a function of a simple form that provides an upper bound on the \peter{information} complexity of a problem.  Some examples of tractability functions commonly considered are provided in Table \ref{tab:commonT} \peter{[I think we can remove the latter sentence, as it is repeated below.]}.
\kachi{}
\begin{table} 
    \caption{Common forms of the tractability function, $T$. \peter{[PETER: I changed the notation below, using ``algebraic" and ``exponential" first, to be in line with the literature.]}}
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		\text{\peter{algebraic strong polynomial}} & \max(1,\varepsilon^{-p}) & 1\\
		\text{\peter{algebraic polynomial}} & \max(1,\varepsilon^{-p})d^{q} & d^{q}\\
		\text{\peter{exponential strong polynomial}} &  [\max(1,\log(1 + \varepsilon^{-1})]^p & 1\\
		\text{\peter{exponential polynomial}} &
		[\max(1,\log(1 + \varepsilon^{-1}))]^p  d^{q} & d^{q} \\
        \text{\peter{algebraic quasi-polynomial}} & 
        \exp\{p(1+\log(\max(1,\varepsilon^{-1}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}\\
        \text{\peter{exponential quasi-polynomial}} & 
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
\label{tab:commonT}
\end{table}

@@

@@

We first consider the simpler case when the \peter{information} complexity is essentially independent of the dimension, $d$.  The proof also introduces the line of argument used for the case where there is $d$-dependence.

@@

For $\varepsilon$ taking on all positive values, $i(\varepsilon,\vp)$ takes on, \peter{by \eqref{eq:Tlim},} all values greater than or equal to $i(\infty,\vp)$, so
\[
i\le C_{\vp} T(\lambda_{i,d}^{-1},1,\vp)+1, \qquad \forall i \ge i(\infty,\vp), \ d \in \naturals.
\]
This implies via our technical assumption \eqref{eq:ptauassume} that, \peter{for any $\tau>1$,}
\begin{gather*}
 K_{\vp,\tau}\,T (\lambda_{i,d}^{-1},1,\tau \vp) \ge
 [T(\lambda_{i,d}^{-1},1, \vp)]^\tau
 \ge
  \left[\frac{(i-1)}{C_{\vp}}\right]^\tau \qquad \forall i \ge i(\infty,\vp), \ d \in \naturals, \\
 \frac{1}{T (\lambda_{i,d}^{-1},1,\tau \vp)} \le
\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(i-1)^\tau}\qquad \forall i \ge \max\{i(\infty,\vp),2\}, \ d \in \naturals .
\end{gather*}
Summing both sides of this last inequality over $i$ from $i(\infty,\vp) +1 \ge 2$ to $\vinfty$ yields
\begin{equation*}
\sup_{d\in\naturals} \sum_{i=i(\infty,\vp) +1 }^\infty \frac{1}{T (\lambda_{i,d}^{-1},1, \tau \vp)}
 \le  K_{\vp,\tau}\, C_{\vp}^\tau
\sum_{i=2}^\infty \frac{1}{(i-1)^\tau} \\
 \le K_{\vp,\tau}\, C_{\vp}^\tau
\zeta (\tau)  < \infty,
\end{equation*}
where $\zeta$ denotes the Riemann zeta function.
This yields \eqref{eq:strong_tractiff3} with $\vp$ replaced by $\vp' = \tau \vp$ and $L_{\vp'} = i(\infty,\vp) +1$, and so we see the necessity of $\eqref{eq:strong_tractiff3}$.

@@

For $\varepsilon$ taking on all positive values, $i(\varepsilon,d,\vp)$ takes on, \peter{by \eqref{eq:Tlim},} all values greater than or equal to $i(\infty,d,\vp)$, so
\[
i\le C_{\vp}\, T(\varepsilon^{-1},d,\vp)+1 \le C_{\vp} T(\lambda_{i,d}^{-1},d,\vp)+1, \qquad \forall i \ge i(\infty,d,\vp),  \ d \in \naturals.
\]
This implies via our technical assumption \eqref{eq:ptauassume} that for any $\tau > 1$,
\begin{gather*}
	K_{\vp,\tau}\,T (\lambda_{i,d}^{-1},d,\tau \vp) \ge
	[T(\lambda_{i,d}^{-1},d, \vp)]^\tau
	\ge
	\left[\frac{(i-1)}{C_{\vp}}\right]^\tau \qquad \forall i \ge i(\infty,d,\vp),  \ d \in \naturals, \\
	 \frac{1}{T (\lambda_{i,d}^{-1},d,\tau \vp)} \le
	\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(i-1)^\tau}\qquad \forall i \ge \max\{i(\infty,d,\vp),2\},  \ d \in \naturals.
\end{gather*}

@@

The \peter{concept of} (strong) polynomial tractability for both the algebraic case and the exponential case will be first considered as follows:

\begin{example}[\peter{Algebraic strong} polynomial tractability]

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-\peter{\vp}}\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]  
where in this case \peter{$\vp=p$} is a scalar.
Then Theorem \ref{thm_main_strong_tract2} yields the  equivalent condition for \peter{algebraic strong polynomial tractability}:
\[
 \sup_{d\in\naturals} \sum_{i=L_p}^\infty \min\{1,\lambda_{i,d}^p\} < \infty \text{ for some } p \ge 0, L_p \in \naturals.
\]
Furthermore, the optimal $p$ is the infimum of all $p$ for which the above condition holds.
This essentially recovers the result in \cite[Theorem 5.1]{NovWoz08a}.
\end{example}

\begin{example}[Algebraic polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-\peter{\vp}}\}\, d^q
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where in this case $\vp = (p,q)$.
Then Theorem \ref{thm_main_tract2} yields the  equivalent condition for algebraic polynomial tractability:
\[
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty \min\{1,\lambda_{i,d}^p\}  < \infty \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
Furthermore, the set of optimal $(p,q)$ is a subset of
\begin{multline*}
\{(p^*,q^*) : \text{the above condition holds for all }(p,q) \in (p^*,\infty) \times (q^*,\infty) \\
\text{but not for any }(p,q) \in [0,p^*/2) \times [0,q^*/2)\}.
\end{multline*}
\peter{[Is the latter statement still correct?]}
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NovWoz08a}. %\fred{same theorem as above??, what about optimality in \cite[Theorem 5.1]{NovWoz08a}}.
\end{example}

\begin{example}[\peter{Exponential strong} polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= [\max\{1,\log(1+\varepsilon^{-1})\}]^{\peter{\vp}}, \quad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]
where \peter{$\vp=p$} is a scalar. Then Theorem \ref{thm_main_strong_tract2} yields the equivalent condition for \peter{exponential strong polynomial tractability:} 
\begin{equation}\label{eq:exp_spt}
 \sup_{d\in\naturals} \sum_{i=L_{p} }^\infty \frac{1}{[\max\{1,\log(1+\lambda_{i,d}^{-1})\}]^{p}}  < \infty,\quad  \text{ for some } p \ge 0, L_{p} \in \naturals.
\end{equation}
In this case the optimal $p$ is the infimum of all $p$ for which the above condition holds.
\end{example}
\begin{example}[Exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)=\left[\max\{1,\log(1+\varepsilon^{-1})\}\right]^p\, d^q, \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where $\vp = (p,q)$. Then Theorem \ref{thm_main_tract2} yields  the equivalent condition for exponential polynomial tractability:
\begin{equation}\label{eq:exp_pt}
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty \frac{1}{[\max\{1,\log(1+\lambda_{i,d}^{-1})\}]^p} < \infty  \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\end{equation}
\end{example}

@@

We say that a problem is $T$-separable tractable \peter{[PETER: I am wondering if ``$T$-separably tractable'' is better? But obviously, you know that better than me.]} with parameter $\vp$ iff there exists a positive constant $C_{\vp}$ which does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq  C_{\vp} \,  \phi(\varepsilon^{-1},p)\, \psi(d,q)\qquad \forall \varepsilon >0, \, d\in\naturals.
\]
A problem is strongly $T$-separable tractable with parameter $\vp$ iff the complexity is independent of $d$, that is, there exists a positive constant $C_{\vp}$ which again does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq C_{\vp}\,\phi(\varepsilon^{-1},p)\qquad \forall \varepsilon >0, \, d\in\naturals,
\] where in this case $\vp = p$ is a scalar. 
Furthermore, we assume the following, \peter{which corresponds to the conditions} in \eqref{eq:Tlim}, \eqref{eq:T0}, and \eqref{eq:ptauassume}. We assume that
\[
\lim_{\varepsilon\rightarrow0}\phi(\varepsilon^{-1},p) = \infty \qquad \forall d\in\naturals, p >0,
\] 
since there are an infinite number of positive eigenvalues. Furthermore, we have that $\phi(\cdot,p)$ is non-decreasing hence we define the following limit
\[ 
\lim_{\varepsilon \rightarrow \infty}\phi(\varepsilon^{-1},p) \peter{>0.}
\] 
Additionally, we have the following technical assumptions. Suppose that, for any real $\tau_1, \tau_2>0$ and any $p,q>0$, we have that 
\begin{align*}
     [\phi(\varepsilon^{-1},p)]^{\tau_1} \le \hat{K}_{p,\tau_1} \phi (\varepsilon^{-1},\tau_1 p)
     \quad \mbox{and} \quad
      [\psi(d,q)]^{\tau_2} \le \hat{K}_{q,\tau_2} \psi (d,\tau_2 q) \qquad \forall \, \varepsilon \in (0,\infty),\, d\in \naturals,
\end{align*} where $\hat{K}_{p,\tau_1}>0$ is a constant that may depend on $p$ and $\tau_1$, but is independent of $\varepsilon$, and where $\hat{K}_{q,\tau_2}>0$ is a constant that may depend on $q$ and $\tau_2$, but is independent of $d$.

Theorem \ref{thm_main_strong_tract2} yields the following equivalent condition for strong $T$-separable tractability:
\begin{equation*} \label{eq:strong_tractiff}
     \sup_{d \in \naturals} \sum_{i = L_p}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty,
\end{equation*}
for some $p>0$ and $L_p \in \naturals$. Furthermore, the optimal $p$ is also the infimum of all $p$ \peter{satisfying the latter condition}.
Theorem \ref{thm_main_tract2} yields the following equivalent condition for separable tractability:
\[
 \sup_{d \in \naturals} \frac{1}{\psi(d,q)} \sum_{i = \lceil L_{(p,q)}\psi(d,q)\rceil}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty,
\] for some $p,q >0$ and \peter{a positive constant $L_{(p,q)}$}.
\end{example}
\kachi{Next, we will introduce another summarized \peter{[what do we mean by ``summarized''?]} notion of tractability. In the following, we will define two functions $\hat{\phi}$ and $\hat{\psi}$ that take the roles of $\varepsilon^{-1}$ and $d$. As will be seen below, this tractability function is equivalent \peter{[[I am not sure it is equivalent.]]} to the so-called \emph{quasi-polynomial tractability}.}
\
\begin{example}[Non-separable tractability]
    Suppose we define the tractability function $T$ by 
\[
 T(\varepsilon^{-1},d,p)= \exp (p \, \hat{\phi}(\varepsilon^{-1})\, \hat{\psi}(d)) = [\exp( \hat{\phi}(\varepsilon^{-1}))]^{p \, \hat{\psi}(d)}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p\in [0,\infty),
\]  with non-decreasing functions $\hat{\phi} : (0,\infty) \to (0,\infty)$ and $\hat{\psi} : \naturals \to (0,\infty)$. 
We say that a problem is non-separable tractable with \peter{scalar} parameter $p$ iff there exists a positive $C_{p}$ such that

>>>>

\title{A Unified Treatment of Tractability of Approximation Problems Defined on Hilbert Spaces}

@@

%TODO: We should check consistency of notation when using $\max$ or $\min$. Sometimes we use ``$(\ldots)$'' for the argument, sometimes we use ``$\{\ldots\}$''  \fred{Fred will do this.}

@@

The information complexity of the problem $\SOL_d$ is given by the function $\comp:(0,\infty) \times \naturals \to \natzero$, where $\comp(\varepsilon,d)$ is the number of linear functionals required by the best admissible algorithm to satisfy the error criterion \eqref{eq:errcrit}.
By definition, the complexity is non-decreasing  as $\varepsilon$ tends to zero.  We also expect it to be non-decreasing as $d$ tends to infinity, since $d$ is typically the number of variables for the functions in $\cf_d$.

For this case where arbitrary linear functionals are allowed, the optimal solution operator is
\[
\APP_d(f,\varepsilon) = \sum_{i=1}^n \SOL_d(u_{i,d}) \ip[\cf_d]{f}{u_i},
\]
and the information complexity of our linear problem is
\begin{equation}\label{eq:comp_def}
\comp(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\},
\end{equation}
see \cite{NovWoz08a}.

Strictly speaking, we are considering the \textit{absolute} error criterion in this paper.
Alternatively, one could also consider the \textit{normalized} error criterion, where the error in \eqref{eq:errcrit} is divided by the so-called \textit{initial error}, which is the error without sampling the function, or, equivalently, the operator norm of $\SOL_d$. In the latter case, it is known (see, e.g., again \cite{NovWoz08a}) that the information complexity is given by
\[
\comp^{\rm norm}(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\, \lambda_{1, d}\}.
\]
This means that, essentially, in the tractability analysis below, all criteria would be normalized by the first eigenvalue $\lambda_{1,d}$. For the sake of brevity and clarity of notation, we restrict ourselves to the absolute error criterion here, but presumably analogous results hold for the normalized setting.

The question of tractability is one of determining how fast the information complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  We want to bound the complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding. We would like to point out that generalized tractability specified by a function $T$ has been considered before in the literature. In particular, there are several papers by Gnewuch and Wo\'{z}niakowski (see \cite{GW06,GW08,GW09,GneWoz11a} ), and also \cite[Chapter 8]{NovWoz08a} is devoted to this topic. To our best knowledge, however, these references mainly focus on the special case of tensor product problems, which we do not do here, and the conditions for tractability derived there are mostly in terms of properties of the function $T$, and not in terms of sums of (functions of) the eigenvalues of a problem. Hence, we consider the present paper an addition to what is presently known on this subject.

@@

\kachi{We denote the set of positive integers by $\naturals$ and and the set of non-negative integers by $\naturals_0$.}
In Section \ref{sec:spt}, we give the definitions of the considered tractability notions, in particular (strong) tractability. We consider strong tractability when the information complexity is independent of the dimension, $d$. We then turn to tractability  when the information complexity may depend on both $d$ and $\varepsilon^{-1}$. In Section \ref{sec:examples} we provide examples of tractability functions for both the algebraic and exponential cases. We also consider other notions of tractability such as quasi-polynomial tractability, and a generalized notion of weak tractability. In the case of generalized weak tractability, we give the necessary and sufficient conditions on the notion of exponential tractability in Section \ref{sec:subh} \peter{[I think I do not quite understand what we mean by the last sentence.]}.

@@

\subsection{Notation and fundamental definitions} \label{sec:not}
Let $T$ be a function given in \eqref{eq:Tspec}.
A tractability function $T$ is a function of a simple form that provides an upper bound on the information complexity of a problem.

@@

\begin{table}
    \caption{Common forms of the tractability function, $T$.}
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		\text{\peter{algebraic strong polynomial}} & \max\{1,\varepsilon^{-p}\} & 1\\
		\text{\peter{algebraic polynomial}} & \max\{1,\varepsilon^{-p}\}d^{q} & d^{q}\\
		\text{\peter{exponential strong polynomial}} &  [\max\{1,\log(1 + \varepsilon^{-1}\}]^p & 1\\
		\text{\peter{exponential polynomial}} &
		[\max\{1,\log(1 + \varepsilon^{-1})\}]^p  d^{q} & d^{q} \\
        \text{\peter{algebraic quasi-polynomial}} &
        \exp\{p(1+\log(\max\{1,\log(1 + \varepsilon^{-1})\})(1+\log(d))\} &
        \exp\{p(1+\log(d))\}\\
        \text{\peter{exponential quasi-polynomial}} &
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
\label{tab:commonT}
\end{table}

@@

We first consider the simpler case when the information complexity is essentially independent of the dimension, $d$.  The proof also introduces the line of argument used for the case where there is $d$-dependence.

@@

For $\varepsilon$ taking on all positive values, $i(\varepsilon,\vp)$ takes on, by \eqref{eq:Tlim}, all values greater than or equal to $i(\infty,\vp)$, so
\[
i\le C_{\vp} T(\lambda_{i,d}^{-1},1,\vp)+1, \qquad \forall i \ge i(\infty,\vp), \ d \in \naturals.
\]
This implies via our technical assumption \eqref{eq:ptauassume} that, for any $\tau>1$,
\begin{gather*}
 K_{\vp,\tau}\,T (\lambda_{i,d}^{-1},1,\tau \vp) \ge
 [T(\lambda_{i,d}^{-1},1, \vp)]^\tau
 \ge
  \left[\frac{(i-1)}{C_{\vp}}\right]^\tau \qquad \forall i \ge i(\infty,\vp), \ d \in \naturals, \\
 \frac{1}{T (\lambda_{i,d}^{-1},1,\tau \vp)} \le
\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(i-1)^\tau}\qquad \forall i \ge \max\{i(\infty,\vp),2\}, \ d \in \naturals .
\end{gather*}
Summing both sides of this last inequality over $i$ from $i(\infty,\vp) +1 \ge 2$ to $\vinfty$ yields
\begin{equation*}
\sup_{d\in\naturals} \sum_{i=i(\infty,\vp) +1 }^\infty \frac{1}{T (\lambda_{i,d}^{-1},1, \tau \vp)}
 \le  K_{\vp,\tau}\, C_{\vp}^\tau
\sum_{i=2}^\infty \frac{1}{(i-1)^\tau} \\
 \le K_{\vp,\tau}\, C_{\vp}^\tau
\zeta (\tau)  < \infty,
\end{equation*}
where $\zeta$ denotes the Riemann zeta function.
This yields \eqref{eq:strong_tractiff3} with $\vp$ replaced by $\vp' = \tau \vp$ and $L_{\vp'} = i(\infty,\vp) +1$, and so we see the necessity of $\eqref{eq:strong_tractiff3}$.

@@

For $\varepsilon$ taking on all positive values, $i(\varepsilon,d,\vp)$ takes on, by \eqref{eq:Tlim}, all values greater than or equal to $i(\infty,d,\vp)$, so
\[
i\le C_{\vp}\, T(\varepsilon^{-1},d,\vp)+1 \le C_{\vp} T(\lambda_{i,d}^{-1},d,\vp)+1, \qquad \forall i \ge i(\infty,d,\vp),  \ d \in \naturals.
\]
This implies via our technical assumption \eqref{eq:ptauassume} that for any $\tau > 1$,
\begin{gather*}
	K_{\vp,\tau}\,T (\lambda_{i,d}^{-1},d,\tau \vp) \ge
	[T(\lambda_{i,d}^{-1},d, \vp)]^\tau
	\ge
	\left[\frac{(i-1)}{C_{\vp}}\right]^\tau \qquad \forall i \ge i(\infty,d,\vp),  \ d \in \naturals, \\
	 \frac{1}{T (\lambda_{i,d}^{-1},d,\tau \vp)} \le
	\frac{K_{\vp,\tau}\, C_{\vp}^\tau}{(i-1)^\tau}\qquad \forall i \ge \max\{i(\infty,d,\vp),2\},  \ d \in \naturals.
\end{gather*}

@@

The concept of (strong) polynomial tractability for both the algebraic case and the exponential case will be first considered as follows:

\begin{example}[Algebraic strong polynomial tractability]

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-\vp}\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]
where in this case $\vp=p$ is a scalar.
Then Theorem \ref{thm_main_strong_tract2} yields the  equivalent condition for algebraic strong polynomial tractability:
\[
 \sup_{d\in\naturals} \sum_{i=L_p}^\infty \min\{1,\lambda_{i,d}^p\} < \infty \text{ for some } p \ge 0, L_p \in \naturals.
\]
Furthermore, the optimal $p$ is the infimum of all $p$ for which the above condition holds.
This essentially recovers the result in \cite[Theorem 5.1]{NovWoz08a}.
\end{example}

\begin{example}[Algebraic polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-\vp}\}\, d^q
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2,
\] where in this case $\vp = (p,q)$.
Then Theorem \ref{thm_main_tract2} yields the  equivalent condition for algebraic polynomial tractability:
\[
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty \min\{1,\lambda_{i,d}^p\}  < \infty \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
Furthermore, the set of optimal $(p,q)$ is a subset of
\begin{multline*}
\{(p^*,q^*) : \text{the above condition holds for all }(p,q) \in (p^*,\infty) \times (q^*,\infty) \\
\text{but not for any }(p,q) \in [0,p^*/2) \times [0,q^*/2)\}.
\end{multline*}
\peter{[Is the latter statement still correct?]}
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NovWoz08a}. %\fred{same theorem as above??, what about optimality in \cite[Theorem 5.1]{NovWoz08a}}.
\end{example}

\begin{example}[Exponential strong polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= [\max\{1,\log(1+\varepsilon^{-1})\}]^{\vp}, \quad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]
where $\vp=p$ is a scalar. Then Theorem \ref{thm_main_strong_tract2} yields the equivalent condition for exponential strong polynomial tractability:
\begin{equation}\label{eq:exp_spt}
 \sup_{d\in\naturals} \sum_{i=L_{p} }^\infty \frac{1}{[\max\{1,\log(1+\lambda_{i,d}^{-1})\}]^{p}}  < \infty,\quad  \text{ for some } p \ge 0, L_{p} \in \naturals.
\end{equation}
In this case the optimal $p$ is the infimum of all $p$ for which the above condition holds.
\end{example}
\begin{example}[Exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)=\left[\max\{1,\log(1+\varepsilon^{-1})\}\right]^p\, d^q, \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2,
\] where $\vp = (p,q)$. Then Theorem \ref{thm_main_tract2} yields  the equivalent condition for exponential polynomial tractability:
\begin{equation}\label{eq:exp_pt}
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty \frac{1}{[\max\{1,\log(1+\lambda_{i,d}^{-1})\}]^p} < \infty  \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\end{equation}
\end{example}

@@

We say that a problem is $T$-separable tractable \peter{[PETER: I am wondering if ``$T$-separably tractable'' is better? But obviously, you know that better than me.]} with parameter $\vp$ iff there exists a positive constant $C_{\vp}$ which does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq  C_{\vp} \,  \phi(\varepsilon^{-1},p)\, \psi(d,q)\qquad \forall \varepsilon >0, \, d\in\naturals.
\]
A problem is strongly $T$-separable tractable with parameter $\vp$ iff the complexity is independent of $d$, that is, there exists a positive constant $C_{\vp}$ which again does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq C_{\vp}\,\phi(\varepsilon^{-1},p)\qquad \forall \varepsilon >0, \, d\in\naturals,
\] where in this case $\vp = p$ is a scalar.
Furthermore, we assume the following, which corresponds to the conditions in \eqref{eq:Tlim}, \eqref{eq:T0}, and \eqref{eq:ptauassume}. We assume that
\[
\lim_{\varepsilon\rightarrow0}\phi(\varepsilon^{-1},p) = \infty \qquad \forall d\in\naturals, p >0,
\]
since there are an infinite number of positive eigenvalues. Furthermore, we have that $\phi(\cdot,p)$ is non-decreasing hence we define the following limit
\[
\lim_{\varepsilon \rightarrow \infty}\phi(\varepsilon^{-1},p) >0.
\]
Additionally, we have the following technical assumptions. Suppose that, for any real $\tau_1, \tau_2>0$ and any $p,q>0$, we have that
\begin{align*}
     [\phi(\varepsilon^{-1},p)]^{\tau_1} \le \hat{K}_{p,\tau_1} \phi (\varepsilon^{-1},\tau_1 p)
     \quad \mbox{and} \quad
      [\psi(d,q)]^{\tau_2} \le \hat{K}_{q,\tau_2} \psi (d,\tau_2 q) \qquad \forall \, \varepsilon \in (0,\infty),\, d\in \naturals,
\end{align*} where $\hat{K}_{p,\tau_1}>0$ is a constant that may depend on $p$ and $\tau_1$, but is independent of $\varepsilon$, and where $\hat{K}_{q,\tau_2}>0$ is a constant that may depend on $q$ and $\tau_2$, but is independent of $d$.

Theorem \ref{thm_main_strong_tract2} yields the following equivalent condition for strong $T$-separable tractability:
\begin{equation*} \label{eq:strong_tractiff}
     \sup_{d \in \naturals} \sum_{i = L_p}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty,
\end{equation*}
for some $p>0$ and $L_p \in \naturals$. Furthermore, the optimal $p$ is also the infimum of all $p$ satisfying the latter condition.
Theorem \ref{thm_main_tract2} yields the following equivalent condition for separable tractability:
\[
 \sup_{d \in \naturals} \frac{1}{\psi(d,q)} \sum_{i = \lceil L_{(p,q)}\psi(d,q)\rceil}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty,
\] for some $p,q >0$ and a positive constant $L_{(p,q)}$.
\end{example}
\kachi{Next, we will introduce another summarized \peter{[what do we mean by ``summarized''?]} notion of tractability. In the following, we will define two functions $\hat{\phi}$ and $\hat{\psi}$ that take the roles of $\varepsilon^{-1}$ and $d$. As will be seen below, this tractability function is equivalent \peter{[[I am not sure it is equivalent.]]} to the so-called \emph{quasi-polynomial tractability}.}
\
\begin{example}[Non-separable tractability]
    Suppose we define the tractability function $T$ by
\[
 T(\varepsilon^{-1},d,p)= \exp (p \, \hat{\phi}(\varepsilon^{-1})\, \hat{\psi}(d)) = [\exp( \hat{\phi}(\varepsilon^{-1}))]^{p \, \hat{\psi}(d)}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p\in [0,\infty),
\]  with non-decreasing functions $\hat{\phi} : (0,\infty) \to (0,\infty)$ and $\hat{\psi} : \naturals \to (0,\infty)$.
We say that a problem is non-separable tractable with \peter{scalar} parameter $p$ iff there exists a positive $C_{p}$ such that
